---
title: 'Homework Five: Functions'
author: "April Wright"
date: "2022-09-23"
output: html_document
---
## Due date: Sept 30, 5 PM.

This homework concerns functions.

1. Make a function that reads in data and does some sort of manipulation of it. Undergrads, use the surveys dataset. Graduates, use your own data. Any manipulation is fine: selection, filtering, value removal. 

```{r}
ac_data <- read_csv ("../data/ac_field_data.csv")
filter_red <- function(dataset = ac_data) {
red_river <- (Month == 6 & Day <= 10)  %>% 
  return(red_river)
}

filter_sabine <- function(dataset = ac_data) {
  sabine_river <- (Month == 6 & Day <= 23)  %>% 
    return(sabine_river)
}

filter_ouachita <- function(dataset = ac_data) {
  ouachita_river <- ac_data %>%  
    filter(Month == 8) 
    return(ouachita_river)
}

# AMW: I took out some of the extra read_csv statements. Once is fine.

# I've also fixed the filter_Ouachita function if you want to copy that. You were missing a pipe to give the data to the filter
filter_ouachita(ac_data)
```


2. Add a test to this function. How will you know if it worked, and alert a user if it did not?

```{r}
ac_data <- read_csv ("ac_field_data.csv") 
filter_red <- function(dataset = ac_data) {
red_river <- (Month == 6 & Day <= 10)  %>% 
if (red_river(Month > 6 & Day > 10)) {
  return(red_river)
} else {
  print ("Red coordinates don't match!")
    }
}

ac_data <- read_csv ("ac_field_data.csv")
filter_sabine <- function(dataset = ac_data) {
  sabine_river <- (Month == 6 & Day <= 23)  %>% 
    if (sabine_river(Month > 6 & Day > 23 & Day <= 10)) {
    return(sabine_river)
} else {
  print ("Sabine coordinates don't match!")
  }
}

ac_data <- read_csv ("ac_field_data.csv")
filter_ouachita <- function(dataset = ac_data) {
  ouachita_river <- (Month == 8)  %>% 
    if (ouachita_river(Month < 8)){
    return(ouachita_river)
} else {
  print("Ouachita coordinates don't match!")
}
}

# AMW: Return statements look good, but the filter and pipe issues from problem #1 are present here, as well
```

3. Make a function that makes a plot. This can be any type of plot that we covered. Violin, box, scatter. Include the ability to save the plot to a file in the function.

```{r}
ac_plot <- function(dataset = ac_data){
samples_months <- ggplot(data = ac_data, mapping = aes (x = Month)) + geom_histogram()
  return(samples_months)
}

#AMW: Missing function declaration.
```

4. For saving to a file, users of your function may want the ability to choose their file name. Make the filename an argument so that users can do so.

```{r}
# Answer here.
samples_months <- ac_plot(ac_data)
ggsave("/cloud/project/homeworks/samples_months.pdf", samples_months)
```
