---
title: "Homework One"
output: html_document
---
# AmW: Good work, 10/10
## Homework One

### Due Aug. 26th at 5pm. Simply write your answers in and save in the homeworks folder in your Rstudio.

This is going to be a low-stakes effort to get you comfortable with RStudio and RMarkdown. First, you're going to download the RMarkdown Document for this homework. Copy and paste this command into the console of your RStudio instance.

```{r}
#download.file(url = "https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2022/main/vignettes/homework_1.Rmd", destfile = "/cloud/project/homeworks/homework_1.Rmd")
```

Can you find where the file downloaded to? Look at the help entry for `download.file` if you aren't sure.

Open the file and answer these two questions: 

1. (1pt) What does `download.file` do? 

```
Downloads the following file into my RStudio Cloud
```

2. (1pt) Where did your HW1.Rmd file download to? 

```
it was downloaded to my homeworks folder in the directory Cloud > project > homeworks 
```

Next, we're going to try to execute some code in an RMarkdown document. Press the green triangle.

```{r}
animals <- c("cat", "rat", "dog")
weights <- c(15, 1, 120)
animals[c(1,3)]
```

What just happened? Annotate the above code with two comments (each worth one point) explaining the code.

#In line 35, the object 'animals' is assigned a vector with "cat" "rat" and "dog" as the series of identifying characters. Line 36 assigns the values "15" "1" and "120" to the object "weights".

#A subset of the vector for the object 'animals' is then extracted to isolate the 1st and 3rd indices of "cat" and "dog".


We can also embed plots in RMarkdown. The below code should embed a barplot. Not very useful, but let me know if you don't see one show up in your file.

```{r}
barplot(names.arg=animals, height=weights)
```

Did the plot show up OK?

```
Yes

```

## Getting to know you all

For the remainder of the assignment, I need to know three things.

1. What type of data will you be working with for your thesis or research project? For example: matrices of phylogenetic characters, spreadsheets of ecological observations, shapefiles from 3D scanning, etc. If you aren't actively doing research, what kinds of scientific questions do you find interesting?

For the first part of my project I will be comparing different PCR techniques, I am unsure what the data analysis for this will look like, but I'd imagine there will be statistical comparisons of detection efficiency with ample amounts of spreadsheet data and potentially the use of the Geneious software.  The second part of my project will involve landscape features and their influence on the presence/absence of the western sand darter.  I will be using GIS, ideally ArcGIS, where I will be working with shapefiles,  CSV/Excel files, and potentially feature classes and rasters if I can find a way to access ArcPro. :)

# AMW: Excellent. I'll see if I can make some time for shapefiles.

2. What are you hoping to get out of class?

I am hoping to get as familiar as possible with R and become as proficient as possible so I can get a job when I leave here, and also so I can have fun learning a new skill.

3. Are there topics that aren't on the schedule that you'd be interested to see? 

I think maps are sick and I get really excited every time I see one so I'm looking forward to the spatial data lesson and would be interested in learning more map things at any time.
