---
title: "Homework Three"
author: "April Wright"
date: "10/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Plotting and choices

## 10 points, due Sept. 16 at 5 PM

```{r}
download.file(url = "https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2022/main/vignettes/homework_3.Rmd", destfile = "/cloud/project/homeworks/homework_3.Rmd")
```

0. Read in the surveys.csv dataset. Remove NAs from the datasets.

1. In this homework, we're going to explore how different ways of visualizing the same data. First, take a look at `geom_col()`. Group the surveys data by the species_id column and count how many of each species there are. Use this count data to make a bar plot of the counts per species. 

```{r}
#Answer here
setwd("/cloud/project/homeworks")
Surveys <- read_csv("surveys.csv")
surveys <- na.omit(Surveys)

```

```{r}
species_grouped <- surveys %>% 
group_by (species_id) %>% 
count (species_id)
```

```{r}
hw_plot <- ggplot(data = species_grouped, mapping = aes(x = species_id, y = n)) + geom_col()
ggsave("/cloud/project/output/hw_plot.pdf", hw_plot, width = 15, height = 10, dpi = 300)

# AMW: Good width to height ratio on this. Looks very clean.
```
2. Take a look at the argument `fct_reorder`. It reorders variables on one or both axes. Try using this to plot the species in order from most to least members.

```{r}
# Answer here

#install the required package
install.packages("forcats")
```

```{r}
library(forcats)
hw_plot_desc <- species_grouped %>%
 ggplot(data = species_grouped, mapping = aes(x = species_id, y=n)) + geom_col(aes(fct_reorder(species_id, n, .desc =TRUE))) 

hw_plot_desc
#ggsave("/cloud/project/output/hw_plot_desc.pdf", hw_plot_desc, width = 15, height = 10, dpi = 300)

```


3. What we canonically think of as a bar plot can also be made in R. These can have some interesting properties, such as being able to fill in bars by other aesthetics. Using the surveys dataset, try to plot the number of members of each species, with the bar filled in by sex.

```{r, eval=FALSE}
mem_species <- ggplot(data = surveys, mapping = aes(x = species_id, fill=sex)) +
  geom_bar()
ggsave("/cloud/project/output/mem_species.pdf", mem_species, width = 15, height = 10, dpi = 300)

```


4. `geom_bar` accepts various arguments. Try `postion="dodge"` or position="stack". How does this change the plot and how you interpret it? 


```{r}
position_dodge <- ggplot(data = surveys, mapping = aes(x = species_id, fill=sex)) +
  geom_bar(position = "dodge")
ggsave("/cloud/project/output/position_dodge.pdf", position_dodge, width = 15, height = 10, dpi = 300)

```
#This changes the visualization of the plot by positioning the two sexes, male and female, side by side on the x axis, rather than positioning them on the same bar and merely differentiating by color.  This would enable an easy side by side comparison of the sex ratio for each species.
```


5. Even a simple histogram can lead to different interpretations. Make a histogram of hindfoot_lengths in the surveys dataset. Try choosing different binwidths. How does a high bindwith (like 100) this change your interpretation of how the data are distributed? 

```{r}
histogram <- ggplot(data = surveys, mapping = aes (x = hindfoot_length)) + geom_histogram(bindwith=30)
ggsave("/cloud/project/output/histogram.pdf", histogram, width = 15, height = 10, dpi = 300)
``` 


6. Next, let's take a look at density plots. First, look at the help for `geom_density`. Density plots are like a smoothed histogram, mostly used for continuous data. But how density is calculated is done using what is called a kernel. To get a sense for what this means, try different kernel types. Some common ones are "triangular", "rectangular", and "gaussian." Try them out, and put the one you think best represents the data in the answer below.


```{r, eval=FALSE}
surveys_gaus <- ggplot(data = surveys, aes(hindfoot_length)) +
    geom_density(fill = "blue", binwidth = 0.5, kernel ="gaussian")
ggsave("/cloud/project/output/surveys_gaus.pdf", surveys_gaus, width = 15, height = 10, dpi = 300)

# AMW: That looks nice.
```

## Graduate Students:

A common set of journal figure requirements several of you submitted include the following: 

+ PDF or PNG
+ At least 300 DPI
+ One column (80 mm) or two column (160mm) wide.

For each of your plots on the above homework, save figures meeting all requirements in your lastname_directory in a directory called `output.`

As a longer-term goal, you each picked a few figures. See if you can find the data that the authors used to make the figures. One of the best ways to learn to make good figures in your field is to copy from what's already published. 

What data did the author's use? 

```
Water quality data and coordinate data.
```

Is the data you're using for class similar to these data? 


```
Yes
```

Could you process your data to be similar to these data?

```
Yes
```

Finally, in the final project in this class, you'll need to produce an R package with five functions: 
 
+ One or two for data cleaning and/or data manipulation
+ One or two for a statistical test (ANOVA, linear models, other specialized analyses)
+ One or two for plotting
+ One of your choice for funsies.

In next week's homework, you'll be expected to pick and write one. So start thinking about one now.
