---
title: "ProjectOne"
output: html_document
---


# Data Project One

## 100 points
## DUE Sept 16 at 11:59 PM

Below are the questions for the first data practical assignment. This project uses the "FossilAnts.csv" file, located in the data directory for the project. The point value of each question is denoted next to it. A blank cell is below each for your answer; feel free to create more blank cells as needed.

0. (5 pts) Create a directory called `projects`, and in it, a subdirectory called `project_one`. Use `download.files` to get [this file]("https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2022/main/vignettes/ProjectOne.Rmd"), and [this one](https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2021/main/projects/project_one/data/FossilAnts.csv).


1. 5 pts. Import the `tidyverse` package and load the data. The data for this part of the practical is located in the data directory. Save the data in a variable called `project_dat`. Print the data to the screen to ensure it loaded correctly.

```{r}
library(tidyverse)
  setwd("/cloud/project/projects/project_one/")
  project_data <- read.csv(file = "FossilAnts.csv")
  project_data
```

2. (5pts) Check the datatypes of each column. There is a column called reference number. This is a static identifier - it should not be changed, and is an indexer used to identify this specimen uniquely. Do we want to treat it as an integer (this is an opinion question - answers may vary).

```
I think with this dataset it would be best for this column to be identified as a character because the reference numbers are very high and could be mistaken as an outlier in a graph if they were accidentally plotted.
```

3. (5 pts) Change the datatype of the reference number column to character. Take a peek at the function as.character().

```{r}
ref_character <-  as.character(project_data$ref_number)

```

4. (5 pts) Look at your data. What are the missing data values? In particular, have a look at the `Tribe` column. In your opinion, are these intelligent missing values for the dataset? Why or why not? If not, how would you like to change them? 

```
These are not intelligent missing values because there are cells that specify as "None" and also empty cells that should presumably  be specified as "NA".  Changing these cells could be done by filling the empty cells in the tribe column with the specification of "NA".   

```

6. (5 points) Are there any columns in your dataset that contain two pieces of data? If so, identify them.

```
# Yes, the taxon column has species listed in some cells, and the full binomical name in others, which includes the genus and species.
```

7. (5pts) Please look at the help page for the `separate` function.

```{r}
# ?separate 
```


7. (10 pts) How could you separate one column into two? 

```{r}
library(dplyr)
 proj_separated <- separate(project_data, taxon, c("genus", "species"))
```

Next, we will test a hypothesis. Your hypthesis is that there are more specimens in the 75 million years ago (mya) - 100 mya interval than the 30 mya to the present interval.

8. (5 pts) Write out the steps you would take to address this question. Will you need to split up the data? Will you need to group the data based on the values in some column? 

#First the old species would be assigned to a variable and then filtered based on years greater than or equal to 75mya and less than or equal to 100mya.  The genus, species, min_ma, and max_ma columns would then be selected to subset them into the requested relevant information.  A similar code would be required for the new species, with the numbers being less than 30mya and less than or equal to 30mya.


9. (15 pts) Perform the operations you described in (8). 

```{r}
old_species <- proj_separated %>%
filter(min_ma >= 75, max_ma <= 100) %>%
select(Genus,species,min_ma, max_ma)

new_species <- proj_separated %>%
filter(min_ma < 30, max_ma <= 30) %>% 
select(Genus,species,min_ma, max_ma)
```

10. (5 pts) Do the results of your code support the hypothesis? 

```
#Answer here

# AMW: Answer? Here?
```


11. (10 pts) Save the dataframe with the split columns into a new directory called `project_one_data_output`. Save it as a csv file called "column_separated.csv" 

```{r}
# Answer here
write.csv(proj_separated, file = "/cloud/project/projects/project_one_data_output/column_separated.csv")
```

12. (10 pts) Produce a histogram of the number of specimens by minimum age. Look at the geom_histogram() help.

```{r}
#Answer here
min_histogram <- ggplot(data = project_data, mapping = aes (x = min_ma)) + geom_histogram(binwidth = 2)
```

13. (5 pts) Change the font size on your histogram so that it can be read comfortably on your computer screen from five feet away. Save the file as "large_font.pdf".

```{r}
min_histogram <- ggplot(data = project_data, mapping = aes (x = min_ma)) + geom_histogram(binwidth = 2) + labs(title = "Number of Specimen by Minimum Age", x="Age", y="Number of Specimen") + theme (text = element_text(size=17.5))
```

14. Finally, produce a histogram of counts for each subfamily.

```{r}
hist_subfam <- ggplot(data = proj_separated, mapping = aes (x = subfamily)) + geom_histogram(stat = "count") + labs(title = "Subfamily Counts", x="Subfamily", y="Counts") + theme(text=element_text(size=13.5)) + theme(axis.text.x = element_text( angle = 90, hjust=.9, vjust=0.5))
```


# Graduate Students

1. You will do the above steps with the classroom dataset.
2. Next, you will choose three data steps from above to perform with your own data. This could be grouping, mutating, cleaning NA values, visualization, or any combination of the three. 
3. Why did you choose these steps? What function does this serve for your thesis and/or other research projects? 

#To be honest, I read question 1 and thought it meant to say my dataset.  Therefore, I redid all of Project One and applied it to my own data.  This can be found in the Project_One_MyData.Rmd file.  I believe it was more than 3 steps.
